---
title: "Inspect and try products without a lake"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Inspect and try products without a lake}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A workspace can show product definitions and existing quality evidence without installing DuckDB or opening a database connection. The IDE package itself contains no compiled code.

```{r}
library(dataraft.ide)
workspace <- new.env(parent = emptyenv())
workspace$orders <- dataraft.core::dr_product(
  "orders", data.frame(amount = c(-10, 20))
) |> dataraft.core::dr_add_quality(~ amount >= 0)
context <- ide_context(workspace)
ide_products(context)
workspace$blocked <- dataraft.core::dr_trial(workspace$orders)
ide_incidents(context)
```

Replace the delivery and explicitly try it again. Trial execution never invokes the configured framework destination writer.

```{r}
workspace$passed <- dataraft.core::dr_trial(
  workspace$orders, data = data.frame(amount = c(10, 20))
)
ide_runs(context)
```

RDS offers a portable local delivery format for the same compiler-free example. Reading this file is an explicit user action, not IDE discovery.

```{r}
path <- tempfile(fileext = ".rds")
saveRDS(data.frame(amount = c(30, 40)), path)
workspace$delivery <- readRDS(path)
ide_profile("binding:delivery", context)
unlink(path)
```

A private file response contains metadata only. Applications must validate the bundled schema, enforce the one MiB bound and check the request ID.

```{r}
private <- tempfile("ide-private-")
dir.create(private, mode = "0700")
response <- file.path(private, "response.json")
ide_emit("quality", response, "vignette", context)
jsonlite::fromJSON(response)$kind
unlink(private, recursive = TRUE)
```

Connected lakes add release, contract, lineage and freshness metadata. Supplying `ide_context(workspace, lake = existing_lake)` never opens a new connection. Reports list frozen report identifiers and creation times, not report values or verification results.
